### 2020-01-23 CalNotes.tex
Corrections N-dimensional meeting AA

Let $\epl > 0$ and let us consider the one-dimensional multiscale stochastic differential equation (SDE)
\begin{equation}\label{eq:SDE_MS}
	\d X_t^\epl = -V'(X_t^\epl) \dd t - \frac1\epl p'\left(\frac{X_t^\epl}\epl\right) + \sqrt{2\sigma} \dd W_t,
\end{equation}
where $\sigma > 0$ and $W_t$ is a standard one-dimensional Brownian motion. The functions $V, p\colon \R \to \R$ are slow and fast potentials driving the dynamics of the solution $X_t^\epl$. Given $N \in \N_{>0}$, we assume the slow potential to be of the form
\begin{equation}
	V(x) = \sum_{i=1}^N \alpha_i V_i(x),
\end{equation}
for coefficients $\alpha_i \in \R$, $i = 1, \ldots, N$, and smooth functions $V_i\colon \R \to \R$. Moreover, we assume $p$ to be smooth and periodic of period $L$. Theory of homogenization \cite{BLP78} guarantees the existence of an SDE of the form
\begin{equation}\label{eq:SDE_HOM}
	\d X_t^0 = - K V'(X) \dd t + \sqrt{2K\sigma} \dd W_t,
\end{equation}
where $W_t$ is the same Brownian motion and where the fast dynamics have been eliminated, such that $X_t^\epl \to X_t^0$ for $\epl\to 0$ in law as random variables in $\mathcal C^0((0, T), \R)$. In the following, we will denote $A_i \defeq K\alpha_i$. The coefficient $K$ is given by the formula
\begin{equation}\label{eq:K_HOM}
	K = \int_0^L (1 + \Phi'(y))^2 \, \mu(\d y),
\end{equation}
with 
\begin{equation}
	\mu(\d y) = Z^{-1} e^{-p(y)/\sigma} \dd y, \quad\text{where}\quad Z = \int_0^L e^{-p(y)/\sigma} \dd y,
\end{equation}
and where the function $\Phi$ is the solution of the elliptic partial differential equation
\begin{equation}
	-p'(y)\Phi'(y) + \sigma \Phi''(y) = p'(y), \quad 0 \leq y \leq L,
\end{equation}
endowed with periodic boundary conditions.

\corr{Failure without pre-processing: explain.}

Let us consider a trajectory $X_{0:T}^\epl \defeq (X_t^\epl, 0 \leq t \leq T)$ coming from the multiscale equation \eqref{eq:SDE_MS}, and the corresponding posterior distribution over the parameter $A$, which we denote by $\mu^\epl(A \mid X_{0:T}^\epl)$. The following result holds.
\begin{theorem} Under assumption \ref{as:regularity} and if $T = \epl^{-\gamma}$ for $\gamma > 0$, then the posterior distribution $\mu^\epl(A \mid X_{0:T}^\epl) = \mathcal N(\bar A^\epl_T,  \sigma^2_T)$ satisfies
	\begin{equation}
	\begin{aligned}
	&\lim_{\epl \to 0} \bar A^\epl_T = \alpha, \\
	&\lim_{\epl \to 0} \sigma^2_T = 0.
	\end{aligned}
	\end{equation}
\end{theorem}
\begin{proof} Proposition \ref{prop:equiv} guarantees that 
	\begin{equation}
	\lim_{\epl \to 0} \abs{\bar A_T - \hat A(X^\epl_{0:T})} = 0,
	\end{equation}
	and that $\sigma^2_T \to 0$ for $\epl \to 0$. Moreover, \cite[Theorem 3.4]{PaS07} yields
	\begin{equation}
	\lim_{\epl \to 0} \hat A(X^\epl_{0:T}) = \alpha,
	\end{equation}
	which completes the proof.
\end{proof}

@@

Let us denote by $A \in \R^N$ the vector of the coefficients $A_i \defeq K \alpha_i$ appearing in the drift term of the homogenized SDE \eqref{eq:SDE_HOM}. In a Bayesian setting, our goal is to determine the posterior distribution $\mu(A \mid X_{0:T})$ given a continuous trajectory $X_{0:T} \defeq (X_t, 0 \leq t \leq T)$. Choosing a Gaussian prior $\mu_0 = \mathcal N(A_0, C_0)$ on $A$, where $A_0 \in \R^N$ and $C_0 \in \R^{N\times N}$ is symmetric positive definite, the posterior distribution admits a density $p(A \mid X_{0:T})$ with respect to the Lebesgue measure which satisfies
\begin{equation}
	p(A \mid X_{0:T}) = \frac1Z \, p(X_{0:T} \mid A) \, p_0(A),
\end{equation}
where $Z$ is the normalization constant, $p_0$ is the density of $\mu_0$, and where Girsanov's formula allows to write the likelihood as
\begin{equation}
	p(X_{0:T} \mid A) = \exp\left\{-\int_0^T \langle A, \mathbf V'(X_t)\rangle \dd X_t - \frac12 \int_0^T \langle A, \mathbf V'(X_t) \rangle^2 \dd t \right\}.
\end{equation}
In the formula above, we denote by $\mathbf V \colon \R \to \R^N$ the vector-valued function whose $i$-th component is $\mathbf V_i = V_i$, and by $\mathbf V'$ the vector of its derivatives, i.e., the Jacobian matrix. Moreover, we denote by $\langle \cdot, \cdot \rangle$ the Euclidean scalar product in $\R^N$ and by $\norm{\cdot}_2$ the corresponding norm. The log-posterior density is then given by
\begin{equation}
\begin{aligned}
	\log p(A \mid X_{0:T}) = &-\log Z -\int_0^T \langle A, \mathbf V'(X_t)\rangle \dd X_t - \frac12 \int_0^T \langle A, \mathbf V'(X_t) \rangle^2 \dd t \\
	&- \frac12 \norm{C_0^{-1/2}(A - A_0)}_2^2.
\end{aligned}
\end{equation}
Since the log-posterior density is quadratic in $A$, the posterior is Gaussian, and it is therefore sufficient to determine its mean and covariance to fully characterize it. We denote by $m_T$ and $C_T$ the mean and covariance matrix, respectively. Let us consider the matrix-valued function $M \colon \R \to \R^{N\times N}$ whose entries are given by
\begin{equation}
	M_{ij} = \frac1T \int_0^T V_i'(X_t) \, V_j'(X_t) \dd t, \quad i, j = 1, \ldots, N,
\end{equation}
and the vector-valued function $h \colon \R \to \R^N$ defined by
\begin{equation}
	h_i = \frac1T \int_0^T V_i'(X_t) \dd X_t, \quad i = 1, \ldots, N.
\end{equation}
If we employ this notation, we can rewrite the log-posterior density as
\begin{equation}
\begin{aligned}
	\log p(A \mid X_{0:T}) = &-\log Z - T\langle A, h\rangle - \frac{T}{2} \langle A, M A\rangle - \frac12 \langle A - A_0, C_0^{-1}(A-A_0) \rangle.
\end{aligned}
\end{equation}
Completing the squares in the log-posterior density, it is possible to show that the posterior is a Gaussian $\mu(A \mid X_{0:T}) = \mathcal N(\bar A_T, C_T)$, whose precision matrix and mean are formally given by
\begin{equation}
\begin{aligned}
	C_T^{-1} &= C_0^{-1} + T M, \\
	C_T^{-1}\bar A_T &= C_0^{-1}A_0 - T h. 
\end{aligned}
\end{equation}
We are now interested in the limit of the posterior distribution for $T \to \infty$. Let us first define the maximum likelihood estimator (MLE) $\widehat A(X_{0:T})$ of $A$, which is obtained by maximizing the log-likelihood function $\log p(X_{0:T} \mid A)$, and which is hence formally given by
\begin{equation}
	\widehat A(X_{0:T}) = -M^{-1}h.
\end{equation}
We now introduce regularity assumptions on the SDE.
\begin{assumption}\label{as:regularity} The potentials $p$ and $V$ are such that 
	\begin{enumerate}
		\item for all $T > 0$, the symmetric matrix $M$ is positive definite and its minimum eigenvalue satisfies $\lambda_{\min}(M) \geq \bar \lambda > 0$,
		\item 	\corr{other assumptions to be added}
	\end{enumerate}
\end{assumption}

We can now state the main result for asymptotic convergence of the posterior distribution.
\begin{proposition}\label{prop:equiv} Given a continuous stochastic process $X_t$ which is ergodic with invariant measure $\mu^{\infty}$ such that $\E^{\mu^\infty}(V_i'(\cdot)) \leq C$ for some constant $C > 0$ and for all $i = 1, \ldots, N$, the posterior $\mu(A \mid X_{0:T})$ contracts to the limit of $\widehat A(X_{0:T})$ for $T \to \infty$.
%	 for $T$ arbitrarily large and under Assumption \ref{as:regularity}, the posterior mean and covariance satisfy
%	\begin{equation}
%	\begin{aligned}
%		&\lim_{T \to \infty} \norm{m_T - \widehat A(X_{0:T})}_2 = 0, \\
%		&\lim_{T \to \infty} \norm{C_T}_F = 0,
%	\end{aligned}
%	\end{equation}
%	i.e., the posterior tends in distribution to the Dirac delta centred in the limit of the MLE.
\end{proposition}
\begin{proof} Let us first consider the covariance matrix. Hua's identity yields
	\begin{equation}
		C_T = T^{-1} \left(M^{-1} - Q^{-1}\right),
	\end{equation}
	where 
	\begin{equation}
		Q = M + T M C_0 M.
	\end{equation}
	The eigenvalues of $Q$ satisfy for all $i = 1, \ldots, N$
	\begin{equation}
		\lambda_i(Q) = \lambda_i(M) + T \lambda_i(M C_0 M) \geq \bar \lambda + T \lambda_{\min}(C_0),
	\end{equation}
	and therefore we obtain
	\begin{equation}\label{eq:boundQinv}
		\lambda_{\max}(Q^{-1}) \leq \frac{1}{\bar \lambda + T \lambda_{\min}(C_0)}.
	\end{equation}
	Similarly, we have a bound for the maximum eigenvalue of $M^{-1}$ given by
	\begin{equation}\label{eq:boundMinv}
		\lambda_{\max}(M^{-1}) \leq \frac1{\bar{\lambda}},
	\end{equation}
	which implies that
	\begin{equation}
		\trace{C_T} \leq \frac{N}{T} \left(\frac1{\bar{\lambda}} + \frac{1}{\bar \lambda + T \lambda_{\min}(C_0)}\right),
	\end{equation}
	and which therefore allows us to conclude that
	\begin{equation}
		\lim_{T\to\infty} \norm{C_T}_F = 0.
	\end{equation}
	We now consider the mean. Replacing the expression of the maximum likelihood estimator, we obtain
	\begin{equation}
		\norm{m_T - \widehat A(X_{0:T})}_2 = T^{-1}\norm{M^{-1}C_0^{-1}A_0 - Q^{-1}\left(C_0^{-1}A_0 -h \right)}_2.
	\end{equation}
	Let us remark that \eqref{eq:boundQinv} implies $\norm{Q^{-1}}_F \to 0$ for $T \to \infty$ and \eqref{eq:boundMinv} implies $\norm{M^{-1}}_F \leq C$ for some $C > 0$ independently of $T$. Moreover, the ergodic theorem guarantees that $\norm{h}_2 \leq C$ for $T$ sufficiently big, and therefore the Cauchy--Schwarz and the triangle inequalities imply 
	\begin{equation}
		\lim_{T\to\infty} \norm{m_T - \widehat A(X_{0:T})}_2 = 0,
	\end{equation}
	which proves the desired result.
\end{proof}

@@

In the previous section, we have shown that posterior distributions over the drift coefficient of the homogenized equation are asymptotically incorrect if multiscale data are replaced into the expression of the likelihood. Hence, the need of pre-processing the data is highlighted. In particular, we consider a filtering approach. Let $k \colon \R^+ \times \R^+ \to \R$ be a kernel function and consider the process $Z^\epl \defeq (Z^{\epl,k}_t, 0 \leq t \leq T)$ defined by the weighted average
\begin{equation}\label{eq:ZDef}
	Z^{\epl,k}_t \defeq \int_0^t k^\epl(t-s)X^\epl_s \dd s,
\end{equation}
where the equality is component-wise in $\R^d$. Let $\beta, \delta > 0$ and let us consider a family of exponential kernel functions defined as
\begin{equation}
	k^\epl(t-s) = C_\beta \delta^{-1/\beta} e^{-(t-s)^\beta/\delta},
\end{equation}
where $C_{\beta}$ is a normalizing constant given by
\begin{equation}
	C_\beta = \beta \, \Gamma(1/\beta)^{-1},
\end{equation}
and where $\Gamma(\cdot)$ is the gamma function. In the following, only in case $\beta = 1$ a rigorous analysis is carried on. Nonetheless, numerical experiments show that for higher values of $\beta$ the performances of estimators computed employing the filter are more robust and qualitatively better.

@@

In the following, we consider the filter with $\beta = 1$, i.e.,
\begin{equation}
	k(t,s) = \frac{1}{\delta} e^{-\frac{t-s}{\delta}},
\end{equation}
since in this case it holds
\begin{equation}
	\d Z^\epl_t = k(0) X^\epl_t \dd t + \int_0^t \partial_t k(t-s) X^\epl_s \dd s \dd t = \frac{1}{\delta} \left ( X^\epl_t - Z^\epl_t \right ) \dd t.
\end{equation}

We will treat analytically two different cases. First, the case where the state $X_t^\epl$ is a vector of $\R^d$ and we wish to find the

Considering the processes $X^\epl$ and $Z^\epl$ together, we obtain a system of stochastic differential equations
\begin{equation}
\label{systemSDE}
\begin{aligned}
\d X^\epl_t &= -\left( \alpha \nabla_x V(X^\epl_t) - \frac{1}\epl \nabla_x p \left ( \frac{X^\epl_t}\epl \right ) \right ) \dd t + \sqrt{2\sigma} \dd W_t, \\
\d Z^\epl_t &= \frac{1}{\delta} \left ( X^\epl_t - Z^\epl_t \right ) \dd t.
\end{aligned}
\end{equation}
Let $\rho^\epl = \rho^\epl(x,z)$ be the stationary distribution of the joint process $(X^\epl, Z^\epl)^\top$. Then $\rho^\epl$ is the solution of the $2d$-dimensional stationary Fokker--Planck equation
\begin{equation}
\begin{aligned}
	\label{eq:FPsystem}
	\sigma \Delta_x \rho^\epl(x,z) & + \left ( \alpha  \Delta_x V(x) + \frac{1}{\epl^2} \Delta_x p \left( \frac{x}\epl \right) + \frac1\delta \right) \rho^\epl(x,z) \\
	& + \left ( \alpha \nabla_x V(x) + \frac{1}\epl \nabla_x p \left ( \frac{x}\epl \right ) \right ) \cdot \nabla_x \rho^\epl(x,z) + \frac{1}{\delta}(z - x) \cdot \nabla_z \rho^\epl(x,z) = 0.
\end{aligned}
\end{equation}

\begin{example}\label{ex:OrnUhl} A closed form solution of \eqref{eq:FPsystem} can be obtained in a simple homogenized case. Let $p(y) = 0$ and the parameters $\alpha$, $\sigma$ be replaced respectively by $A$ and $\Sigma$. Then, if $V(x) = x^2/2$, equation \eqref{eq:FPsystem} has the analytical solution 
\begin{equation}
\rho^0(x,z) = \frac{1}{A_{\rho^0}} \exp\left(-\frac{A}{\Sigma} \frac{x^2}{2} - \frac{1}{\delta \Sigma} \frac{(x - (1+A \delta)z)^2}{2}\right),
\end{equation}
where
\begin{equation}
A_{\rho^0} = \int_{\R} \int_{\R} \exp\left(-\frac{A}{\Sigma} \frac{x^2}{2} - \frac{1}{\delta \Sigma} \frac{(x - (1+A \delta)z)^2}{2}\right) \dd x \dd z.
\end{equation}
This is the density of a multivariate normal distribution $\mathcal N(0, \Gamma)$, where the covariance matrix is given by
\begin{equation}
\Gamma = \frac{\Sigma}{A (1 + A\delta)} \begin{pmatrix} 1+A\delta & 1 \\ 1 & 1 \end{pmatrix}.
\end{equation}
Let us remark that this distribution can be obtained from direct computations involving Gaussian processes. In particular, it is known that $X_t \sim \mathcal{GP}(\mu_t, \mathcal C(t, s))$, where 
\begin{equation}
	\mu_t = X_0 e^{-At}, \quad \mathcal C(t, s) = \frac{\Sigma}{A} \left( e^{-A|t-s|} - e^{-A(t+s)} \right).
\end{equation}
The basic properties of Gaussian processes imply that $Z_t$ is a Gaussian process, and that the couple $(X_t, Z_t)^\top$ is a Gaussian process, too, whose mean and covariance are computable explicitly.
\end{example}

@@

\sigma \Delta_x \phi^\epl(x) + \left ( \alpha \Delta_x V(x) + \frac1{\epl^2} \Delta_x p \left( \frac{x}\epl \right) \right) \phi^\epl(x) + \left ( \alpha \nabla_xV(x) + \frac{1}\epl \nabla_xp \left ( \frac{x}\epl \right ) \right ) \cdot \nabla_x\phi^\epl(x) = 0,

@@

\phi^\epl(x) = \frac{1}{C_{\phi^\epl}} \exp\left(- \frac{\alpha}{\sigma} V(x) - \frac{1}{\sigma} p \left ( \frac{x}\epl \right )\right),

@@

\psi^\epl(z) + z \cdot \nabla_z \psi^\epl(z) = \nabla_z \cdot \int_{\R^d} x \rho^\epl(x,z) \dd x,

@@

	\nabla_z \cdot (z \psi^\epl(z)) = \nabla_z \cdot \int_{\R^d} x \rho^\epl(x,z) \dd x.

@@

z \psi^\epl(z) = \int_{\R^d} x \rho^\epl(x,z) \dd x,

@@

z = \int_{\R^d} x \phi^\epl(x) R^\epl(x,z) \dd x.

>>>>

Let $\epl > 0$ and let us consider the one-dimensional multiscale stochastic differential equation (SDE)
\begin{equation}\label{eq:SDE_MS}
	\d X_t^\epl = -\alpha \cdot V'(X_t^\epl) \dd t - \frac1\epl p'\left(\frac{X_t^\epl}\epl\right) + \sqrt{2\sigma} \dd W_t,
\end{equation}
where $\alpha_i \in \R$, $i = 1, \ldots, N$ and $\sigma > 0$ are the drift and diffusion coefficients respectively and $W_t$ is a standard one-dimensional Brownian motion. The functions $V\colon \R \to \R^N$ and $p\colon \R \to \R$ are slow and fast potentials driving the dynamics of the solution $X_t^\epl$. Given $N \in \N_{>0}$, we assume the slow potential to be of the form
\begin{equation}
	V(x) = \begin{pmatrix} V_1(x) & V_2(x) & \cdots & V_N(x) \end{pmatrix}^\top,
\end{equation}
for smooth functions $V_i\colon \R \to \R$, $i = 1, \ldots, N$. Moreover, we assume $p$ to be smooth and periodic of period $L$. Theory of homogenization \cite{BLP78} guarantees the existence of an SDE of the form
\begin{equation}\label{eq:SDE_HOM}
	\d X_t^0 = - A \cdot V'(X_t^0) \dd t + \sqrt{2\Sigma} \dd W_t,
\end{equation}
where $W_t$ is the same Brownian motion and where the fast dynamics have been eliminated, such that $X_t^\epl \to X_t^0$ for $\epl\to 0$ in law as random variables in $\mathcal C^0((0, T), \R)$. In particular, we have $A_i = K\alpha_i$ and $\Sigma = K \sigma$, where the coefficient $K$ is given by the formula
\begin{equation}\label{eq:K_HOM}
	K = \int_0^L (1 + \Phi'(y))^2 \, \mu(\d y),
\end{equation}
with 
\begin{equation}
	\mu(\d y) = Z^{-1} e^{-p(y)/\sigma} \dd y, \quad\text{where}\quad Z = \int_0^L e^{-p(y)/\sigma} \dd y,
\end{equation}
and where the function $\Phi$ is the solution of the elliptic partial differential equation
\begin{equation}
	-p'(y)\Phi'(y) + \sigma \Phi''(y) = p'(y), \quad 0 \leq y \leq L,
\end{equation}
endowed with periodic boundary conditions.

\subsection{Failure without pre-processing \corr{rewrite}}

Let us consider a trajectory $X^\epl \defeq (X_t^\epl, 0 \leq t \leq T)$ coming from the multiscale equation \eqref{eq:SDE_MS}, and the corresponding posterior distribution over the parameter $A$, which we denote by $\mu^\epl(A \mid X^\epl)$. The following result holds.
\begin{theorem} Under assumption \ref{as:regularity} and if $T = \epl^{-\gamma}$ for $\gamma > 0$, then the posterior distribution $\mu^\epl(A \mid X^\epl) = \mathcal N(\bar A^\epl_T,  \sigma^2_T)$ satisfies
	\begin{equation}
	\begin{aligned}
	&\lim_{\epl \to 0} \bar A^\epl_T = \alpha, \\
	&\lim_{\epl \to 0} \sigma^2_T = 0.
	\end{aligned}
	\end{equation}
\end{theorem}
\begin{proof} Proposition \ref{prop:equiv} guarantees that 
	\begin{equation}
	\lim_{\epl \to 0} \abs{\bar A_T - \hat A(X^\epl_{0:T})} = 0,
	\end{equation}
	and that $\sigma^2_T \to 0$ for $\epl \to 0$. Moreover, \cite[Theorem 3.4]{PaS07} yields
	\begin{equation}
	\lim_{\epl \to 0} \hat A(X^\epl_{0:T}) = \alpha,
	\end{equation}
	which completes the proof.
\end{proof}

@@

In a Bayesian setting, our goal is to determine the posterior distribution $\mu(A \mid X)$ given a continuous trajectory $X \defeq (X_t, 0 \leq t \leq T)$. Choosing a Gaussian prior $\mu_0 = \mathcal N(A_0, C_0)$ on $A$, where $A_0 \in \R^N$ and $C_0 \in \R^{N\times N}$ is symmetric positive definite, the posterior distribution admits a density $p(A \mid X)$ with respect to the Lebesgue measure which satisfies
\begin{equation}
	p(A \mid X) = \frac1Z \, p(X \mid A) \, p_0(A),
\end{equation}
where $Z$ is the normalization constant, $p_0$ is the density of $\mu_0$, and where Girsanov's formula allows to write the likelihood as
\begin{equation}
	p(X \mid A) = \exp\left(-\frac{I(X\mid A)}{2\Sigma} \right), 
\end{equation}
and where
\begin{equation}
	I(X\mid A) = \int_0^T A \cdot V'(X_t) \dd X_t + \frac12 \int_0^T \left( A \cdot V'(X_t) \right)^2 \dd t.
\end{equation}

The log-posterior density is then given by
\begin{equation}
\begin{aligned}
	\log p(A \mid X) = &-\log Z -\frac1{2\Sigma}\int_0^T A \cdot V'(X_t) \dd X_t - \frac1{4\Sigma} \int_0^T \left(A \cdot V'(X_t) \right)^2 \dd t \\
	&- \frac12 \norm{C_0^{-1/2}(A - A_0)}_2^2.
\end{aligned}
\end{equation}
Since the log-posterior density is quadratic in $A$, the posterior is Gaussian, and it is therefore sufficient to determine its mean and covariance to fully characterize it. We denote by $m_T$ and $C_T$ the mean and covariance matrix, respectively. Let us consider the matrix-valued function $M \colon \R \to \R^{N\times N}$ whose entries are given by
\begin{equation}
	M_{ij} = \frac1{2\Sigma T} \int_0^T V_i'(X_t) \, V_j'(X_t) \dd t, \quad i, j = 1, \ldots, N,
\end{equation}
and the vector-valued function $h \colon \R \to \R^N$ defined by
\begin{equation}
	h_i = \frac1{2\Sigma T} \int_0^T V_i'(X_t) \dd X_t, \quad i = 1, \ldots, N.
\end{equation}
Let us remark that we can rewrite $M$ and $h$ in vector notation as
\begin{equation}
		M = \frac1{2\Sigma T} \int_0^T V'(X_t) \otimes V'(X_t) \dd t, \quad h = \frac1{2\Sigma T} \int_0^T V'(X_t) \dd X_t,
\end{equation}
where $\otimes$ denotes the tensor product in $\R^N$. Therefore, we can rewrite the log-posterior density as
\begin{equation}
\begin{aligned}
	\log p(A \mid X) = &-\log Z - T A \cdot h - \frac{T}{2} A \cdot M A - \frac12 (A - A_0) \cdot C_0^{-1}(A-A_0).
\end{aligned}
\end{equation}
Completing the squares in the log-posterior density, it is possible to show that the posterior is a Gaussian $\mu(A \mid X) = \mathcal N(m_T, C_T)$, whose precision matrix and mean are formally given by
\begin{equation}
\begin{aligned}
	C_T^{-1} &= C_0^{-1} + T M, \\
	C_T^{-1}m_T &= C_0^{-1}A_0 - T h. 
\end{aligned}
\end{equation}
We are now interested in the limit of the posterior distribution for $T \to \infty$. Let us first define the maximum likelihood estimator (MLE) $\widehat A(X)$ of $A$, which is obtained by maximizing the log-likelihood function $\log p(X \mid A)$, and which is hence formally given by
\begin{equation}
	\widehat A(X) = -M^{-1}h.
\end{equation}
We now introduce regularity assumptions on the SDE.
\begin{assumption}\label{as:regularity} The potentials $p$ and $V$ are such that 
	\begin{enumerate}
		\item for all $T > 0$, the symmetric matrix $M$ is positive definite and its minimum eigenvalue satisfies $\lambda_{\min}(M) \geq \bar \lambda > 0$,
		\item 	\corr{other assumptions to be added}
	\end{enumerate}
\end{assumption}

We can now state the main result for asymptotic convergence of the posterior distribution.
\begin{proposition}\label{prop:equiv} Given a continuous stochastic process $X \defeq (X_t, 0 \leq t, \leq T)$ which is ergodic with invariant measure $\mu^{\infty}$ such that $\E^{\mu^\infty}(V_i'(\cdot)) \leq C$ for some constant $C > 0$ and for all $i = 1, \ldots, N$, the posterior $\mu(A \mid X)$ contracts to the limit of $\widehat A(X)$ for $T \to \infty$.
%	 for $T$ arbitrarily large and under Assumption \ref{as:regularity}, the posterior mean and covariance satisfy
%	\begin{equation}
%	\begin{aligned}
%		&\lim_{T \to \infty} \norm{m_T - \widehat A(X)}_2 = 0, \\
%		&\lim_{T \to \infty} \norm{C_T}_F = 0,
%	\end{aligned}
%	\end{equation}
%	i.e., the posterior tends in distribution to the Dirac delta centred in the limit of the MLE.
\end{proposition}
\begin{proof} Let us first consider the covariance matrix. Hua's identity yields
	\begin{equation}
		C_T = T^{-1} \left(M^{-1} - Q^{-1}\right),
	\end{equation}
	where 
	\begin{equation}
		Q = M + T M C_0 M.
	\end{equation}
	The eigenvalues of $Q$ satisfy for all $i = 1, \ldots, N$
	\begin{equation}
		\lambda_i(Q) = \lambda_i(M) + T \lambda_i(M C_0 M) \geq \bar \lambda + T \lambda_{\min}(C_0),
	\end{equation}
	and therefore we obtain
	\begin{equation}\label{eq:boundQinv}
		\lambda_{\max}(Q^{-1}) \leq \frac{1}{\bar \lambda + T \lambda_{\min}(C_0)}.
	\end{equation}
	Similarly, we have a bound for the maximum eigenvalue of $M^{-1}$ given by
	\begin{equation}\label{eq:boundMinv}
		\lambda_{\max}(M^{-1}) \leq \frac1{\bar{\lambda}},
	\end{equation}
	which implies that
	\begin{equation}
		\trace{C_T} \leq \frac{N}{T} \left(\frac1{\bar{\lambda}} + \frac{1}{\bar \lambda + T \lambda_{\min}(C_0)}\right),
	\end{equation}
	and which therefore allows us to conclude that
	\begin{equation}
		\lim_{T\to\infty} \norm{C_T}_F = 0.
	\end{equation}
	We now consider the mean. Replacing the expression of the maximum likelihood estimator, we obtain
	\begin{equation}
		\norm{m_T - \widehat A(X)}_2 = T^{-1}\norm{M^{-1}C_0^{-1}A_0 - Q^{-1}\left(C_0^{-1}A_0 - Th \right)}_2.
	\end{equation}
	Let us remark that \eqref{eq:boundQinv} implies $\norm{Q^{-1}}_F \to 0$ for $T \to \infty$ and \eqref{eq:boundMinv} implies $\norm{M^{-1}}_F \leq C$ for some $C > 0$ independently of $T$. Moreover, the ergodic theorem guarantees that $\norm{h}_2 \leq C$ for $T$ sufficiently big, and therefore the Cauchy--Schwarz and the triangle inequalities imply 
	\begin{equation}
		\lim_{T\to\infty} \norm{m_T - \widehat A(X)}_2 = 0,
	\end{equation}
	which proves the desired result.
\end{proof}

@@

%Let us consider $N = 1$ and write $A_1 = A$ and $V_1 = V$, respectively. The posterior distribution of $A$ given a trajectory $(X_t, 0 \leq t \leq T)$ is in this case a Gaussian $\mathcal N(\bar A_T, \sigma^2_T)$ where
%\begin{equation}
%\begin{aligned}
%	\bar A_T &= \frac{A_0}{1+\sigma_0^2\int_0^T V'(X_t)^2 \dd t} - \frac{\sigma_0^2\int_0^T  V'(X_t) \dd X_t}{1+\sigma_0^2\int_0^T V'(X_t)^2 \dd t}, \\
%	\sigma^{-2}_T &= \sigma_0^{-2} + \int_0^T V'(X_t)^2 \dd t.
%\end{aligned}
%\end{equation}
%We will in the following sections analyse the convergence of the posterior, both in case data from the multiscale process is not treated and in case they are pre-processed.

In the previous section, we have shown that posterior distributions over the drift coefficient of the homogenized equation are asymptotically incorrect if multiscale data are replaced into the expression of the likelihood. Hence, the need of pre-processing the data is highlighted. In particular, we consider a filtering approach. Let $k \colon \R^+ \times \R^+ \to \R$ be a kernel function and consider the process $Z^\epl \defeq (Z^{\epl,k}_t, 0 \leq t \leq T)$ defined by the weighted average
\begin{equation}\label{eq:ZDef}
	Z^{\epl,k}_t \defeq \int_0^t k^\epl(t, s)X^\epl_s \dd s.
\end{equation}
Let $\beta, \delta > 0$ and let us consider a family of exponential kernel functions defined as
\begin{equation}
	k^\epl(t, s) = C_\beta \delta^{-1/\beta} e^{-(t-s)^\beta/\delta},
\end{equation}
where $C_{\beta}$ is a normalizing constant given by
\begin{equation}
	C_\beta = \beta \, \Gamma(1/\beta)^{-1},
\end{equation}
and where $\Gamma(\cdot)$ is the gamma function. In the following, only in case $\beta = 1$ a rigorous analysis is carried on. Nonetheless, numerical experiments show that for higher values of $\beta$ the performances of estimators computed employing the filter are more robust and qualitatively better.

@@

In the following, we consider the filter with $\beta = 1$, i.e.,
\begin{equation}
	k(t,s) = \frac{1}{\delta} e^{-\frac{t-s}{\delta}},
\end{equation}
since in this case it holds
\begin{equation}
	\d Z^\epl_t = k(t,t) X^\epl_t \dd t + \int_0^t \partial_t k(t,s) X^\epl_s \dd s \dd t = \frac{1}{\delta} \left ( X^\epl_t - Z^\epl_t \right ) \dd t.
\end{equation}
Considering the processes $X^\epl$ and $Z^\epl$ together, we obtain a system of stochastic differential equations
\begin{equation}
\label{systemSDE}
\begin{aligned}
\d X_t^\epl &= -\alpha \cdot V'(X_t^\epl) \dd t - \frac1\epl p'\left(\frac{X_t^\epl}\epl\right) + \sqrt{2\sigma} \dd W_t, \\
\d Z^\epl_t &= \frac{1}{\delta} \left ( X^\epl_t - Z^\epl_t \right ) \dd t.
\end{aligned}
\end{equation}
Let $\rho^\epl = \rho^\epl(x,z)$ be the stationary distribution of the joint process $(X^\epl, Z^\epl)^\top$. \corr{How do we show that the couple $(X^\epl, Z^\epl)^\top$ is ergodic.} Then $\rho^\epl$ is the solution of the two-dimensional stationary Fokker--Planck equation
\begin{equation}
\begin{aligned}
	\label{eq:FPsystem}
	\sigma \partial^2_{xx} \rho^\epl(x,z) & + \left ( \alpha \cdot V''(x) + \frac{1}{\epl^2} p'' \left ( \frac{x}\epl \right ) + \frac{1}{\delta} \right ) \rho^\epl(x,z) \\
	& + \left ( \alpha \cdot V'(x) + \frac{1}\epl p' \left ( \frac{x}\epl \right ) \right ) \partial_x \rho^\epl(x,z) + \frac{1}{\delta}(z - x) \partial_z \rho^\epl(x,z) = 0.
\end{aligned}
\end{equation}

\begin{example}\label{ex:OrnUhl} A closed form solution of \eqref{eq:FPsystem} can be obtained in a simple homogenized case with the dimension of the parameter $N=1$. Let $p(y) = 0$ and the parameters $\alpha$, $\sigma$ be replaced respectively by $A$ and $\Sigma$. Then, if $V(x) = x^2/2$, equation \eqref{eq:FPsystem} has the analytical solution 
\begin{equation}
\rho^0(x,z) = \frac{1}{C_{\rho^0}} \exp\left(-\frac{A}{\Sigma} \frac{x^2}{2} - \frac{1}{\delta \Sigma} \frac{(x - (1+A \delta)z)^2}{2}\right),
\end{equation}
where
\begin{equation}
C_{\rho^0} = \int_{\R} \int_{\R} \exp\left(-\frac{A}{\Sigma} \frac{x^2}{2} - \frac{1}{\delta \Sigma} \frac{(x - (1+A \delta)z)^2}{2}\right) \dd x \dd z.
\end{equation}
This is the density of a multivariate normal distribution $\mathcal N(0, \Gamma)$, where the covariance matrix is given by
\begin{equation}
\Gamma = \frac{\Sigma}{A (1 + A\delta)} \begin{pmatrix} 1+A\delta & 1 \\ 1 & 1 \end{pmatrix}.
\end{equation}
Let us remark that this distribution can be obtained from direct computations involving Gaussian processes. In particular, it is known that $X_t \sim \mathcal{GP}(\mu_t, \mathcal C(t, s))$, where 
\begin{equation}
	\mu_t = X_0 e^{-At}, \quad \mathcal C(t, s) = \frac{\Sigma}{A} \left( e^{-A|t-s|} - e^{-A(t+s)} \right).
\end{equation}
The basic properties of Gaussian processes imply that $Z_t$ is a Gaussian process, and that the couple $(X_t, Z_t)^\top$ is a Gaussian process, too, whose mean and covariance are computable explicitly.
\end{example}

@@

\sigma (\phi^\epl)''(x) + \left ( \alpha \cdot V''(x) + \frac{1}{\epl^2} p'' \left ( \frac{x}\epl \right ) \right ) \phi^\epl(x) + \left ( \alpha \cdot V'(x) + \frac{1}\epl p' \left ( \frac{x}\epl \right ) \right ) (\phi^\epl)'(x) = 0,

@@

\phi^\epl(x) = \frac{1}{C_{\phi^\epl}} \exp\left(- \frac{1}{\sigma} \alpha \cdot V(x) - \frac{1}{\sigma} p \left ( \frac{x}\epl \right )\right),

@@

\psi^\epl(z) + z (\psi^\epl)'(z) = \frac{\d}{\d z} \int_{\R} x \rho^\epl(x,z) \dd x,

@@

	\frac{\d}{\d z} (z \psi^\epl(z)) = \frac{\d}{\d z} \int_{\R} x \rho^\epl(x,z) \dd x.

@@

z \psi^\epl(z) = \int_{\R} x \rho^\epl(x,z) \dd x,

@@

z = \int_{\R} x \phi^\epl(x) R^\epl(x,z) \dd x.
